Declarative Deployment Reconfiguration of Pulverised Collective Adaptive Systems at Runtime \thanks{ This work has been supported by the Italian PRIN Project COMMON-WEARS (2020HCWWLP) and the EU/MUR FSE REACT-EU PON R\&I 2014-2020. }

### 2023-05-21 paper-2023-acsos-pulverization.tex
fix footnote link
Bring II.C into III.A, + refactor presentation of the DSLs + provide a sentence clarifying the section structure -- @danysk @nicolasfara pls check

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{acronym}
\usepackage{listing-style}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes.geometric,positioning,automata,calc}
\usepackage[hidelinks]{hyperref}
\usepackage[inline]{enumitem}
\usepackage{siunitx}
\DeclareSIUnit{\decibelmeter}{dBm}

@@

\subsection{Middlewares and (Domain-Specific) Languages for Deployed Systems Specification and Execution}

@@

From the previous section,
we can identify two main pieces of information for the definition of a deployment plan for a pulverised system:
\begin{enumerate*}[label=\it{(\roman*)}]
  \item the definition of the \emph{target} deployment domain (\emph{infrastructure}), i.e., the specification of the \emph{requirements} posed on hosts in terms of required (or preferred) capabilities; and
  \item the definition of the \emph{deployment plan}, i.e., the allocation of the pulverised components to the infrastructural elements, together with constraints and reconfiguration policies.
\end{enumerate*}
%
Coherently with the \emph{single responsibility principle},
we capture these two descriptive needs into two distinct \acp{dsl}: the \dslSys{} and the \dslDep{}.
%
An overall view of the approach in terms of the provided tools and the main modelling concepts is provided in \Cref{fig:approach}.

\subsection{The \dslSys{}: Components and Required Capabilities}

@@

\subsection{The \dslDep{}: Deployment Domain, Mapping, and Reconfiguration}

@@

\subsection{Implementation details and underlying runtime}

We implemented the proposed \ac{dsl} in Kotlin.
%
We wanted a language with a strongly typed syntax to help developers with better contextual assist,
and we wanted the possibility to target multiple runtimes
(Kotlin currently targets the JVM, JavaScript, and native code for many platforms\footnote{\url{https://archive.is/JSiTg}},
including mobile and wearable devices)
and be as interoperable as possible with other languages.
%
Scala could have been a good alternative,
but we deemed the multiplatform support of Kotlin a better fit for our needs.
%
The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/nicolasfara/pulverization-framework}}.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{acronym}
\usepackage{listing-style}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes.geometric,positioning,automata,calc}
\usepackage[hidelinks]{hyperref}
\usepackage[inline]{enumitem}
\usepackage{siunitx}
\DeclareSIUnit{\decibelmeter}{dBm}
\providecommand{\decibel}{dB}
\providecommand{\decibelmeter}{dBm}
\providecommand{\qty}[2]{#1#2}

@@

From the previous section,
we can identify three main pieces of information for the definition of a pulverised system and its deployment:
\begin{enumerate}[label=\it{(\roman*)}]
  \item the description of the logical system and its partitioning into components and deployment units;
  \item the definition of the \emph{target} deployment domain (\emph{infrastructure}), i.e., the specification of the \emph{requirements} posed on hosts in terms of required (or preferred) capabilities; and
  \item the definition of the \emph{deployment plan}, i.e., the allocation of the pulverised components to the infrastructural elements, together with constraints and reconfiguration policies.
\end{enumerate}
%
%Coherently with the \emph{single responsibility principle},
We capture these %two 
 descriptive needs into two distinct \acp{dsl}: the \dslSys{} and the \dslDep{}.
%
An overall view of the approach in terms of the provided tools and the main modelling concepts is provided in \Cref{fig:approach}.

In the following,
 we first clarify the approach (\Cref{sec:contrib:general}),
 then cover the two \acp{dsl}
 (\Cref{sec:contrib:dslsys,sec:contrib:dsldep}),
 and finally provide details on the runtime system (\Cref{sec:contrib:runtime}).

\subsection{On Middlewares and (Domain-Specific) Languages for Deployed Systems Specification and Execution}
\label{sec:contrib:general}

@@

\subsection{The \dslSys{}: Components and Required Capabilities}
\label{sec:contrib:dslsys}

@@

\subsection{The \dslDep{}: Deployment Domain, Mapping, and Reconfiguration}
\label{sec:contrib:dsldep}

@@

\subsection{Implementation details and underlying runtime}
\label{sec:contrib:runtime}

We implemented the proposed \ac{dsl} in Kotlin.
%
We wanted a language with a strongly typed syntax to help developers with better contextual assist,
and we wanted the possibility to target multiple runtimes
(Kotlin currently targets the JVM, JavaScript, and native code for many platforms\footnote{\url{https://archive.is/JSiTg}},
including mobile and wearable devices)
and be as interoperable as possible with other languages.
%
Scala could have been a good alternative,
but we deemed the multiplatform support of Kotlin a better fit for our needs.
%
The framework has been open sourced and released with a permissive license\footnote{\url{https://github.com/pulvreakt/pulvreakt}}.
